---
title: "Main Model"
author: "Nick Strayer"
date: "March 26, 2015"
output:
  html_document:
    fig_height: 6
    fig_width: 9
---

```{r echo=FALSE}
library(ggplot2)
library(reshape2)
library(grid)
library(gridExtra)
setwd("/Users/Nick/mysisModeling") #Doesnt actually do anything in chunks.
```
---

## Check for the existance of needed datafiles. 

If the files are not present, then run their corresponding scripts and generate them. 

The corresponding R Markdown scripts associated with these files can be found on RPub: 

  - [Solar Data Cleaning](http://rpubs.com/nstrayer/64339)
  
  - [Thermocline Model](http://rpubs.com/nstrayer/thermoclineModel)

  - [Mysocline Depth Model](http://rpubs.com/nstrayer/64310)

```{r}
setwd("/Users/Nick/mysisModeling") #Locate ourselves in the computer. 

#Check for solar data
if(!file.exists("data/light_day_moon_hour.csv")){ source("dataGen/solarData.r") } 

#Check for thermocline data
if(!file.exists("data/Depth_Thermocline_Hour.csv")){ source("dataGen/thermoclineLevels.r") } 

#Mysocline data
if(!file.exists("data/mysocline_hour.csv")){ source("dataGen/depthModel.r") }

#Mysocline data
if(!file.exists("data/FoodAvail_Hour.csv")){ source("dataGen/foodAvailability.r") }

#Read in the data now. 
depthData = read.csv("data/mysocline_hour.csv")$x
foodCurve = read.csv("data/FoodAvail_hour.csv")
foodAvail = foodCurve$foodAvail
foodVar   = foodCurve$variability
```

## Class and method declarations: 

I have chosen to use the S4 object oriented programing style. The details are minor but more info can be found [here](https://www.stat.auckland.ac.nz/~stat782/downloads/08-Objects.pdf). 

### Set up the mysis class: 

This basically is a digital mysis that simplistically holds a couple of peices of information and upon which the model will act. 

```{r, results='hide'}
setClass("mysis",
         representation(
           energy    = "numeric",    # The energy reserves are a numeric value
           migrating = "logical",    # If mysis is migrating is a logical value
           alive     = "logical",    # Alive still?
           depth     = "numeric"),   
         prototype(
           energy    = 0,            # The default instantiated mysis starts with zero energy...
           migrating = FALSE,        # and not migrating...
           alive     = TRUE,         # alive
           depth     = 100),         # and at the bottom.  
         )

#Set the show method, basically how we want the program to display the info about the mysis on calling. 
setMethod("show", "mysis", 
          function(object){
            if (object@alive){
              print(object@energy)
              print(object@migrating)
              print(object@depth)
            } else {
              print("dead")
            }
          })
```

__Real quick we need to make a model for decisions to migrate based upon condition:__

The form of this model is a logistic curve: 
$$f(x) = \frac{1}{1 + e^{-k(x - m)}}$$
Where __$k$__ is an indicator of steepness, and __$m$__ is the midepoint of the curve.

```{r}
migrationProb = function(condition){ #I am choosing to leave the random roll outside of script for continuity, could be changed
  m = 120 #value of curve midpoint
  k = 0.03 #steepness of curve
  x = condition 
  
  dist = 1/(1 + exp(-k* (x - m)))   
  return(1 - dist)
}
```
 


### The decision tree model: 

This is a stochastic model, meaning it utilizes randomness to simulate the real world. 

The function of the decision tree is such: 

- Draw random number to decide if migrating: 
  - If the number drawn is above the ratio needed to migrate (food ratio) then the mysis migrates. 
    - After migration is decided, another draw is done to see if the mysis is killed by predation. 


```{r, results='hide'}

rewardUnits = .67
migrationCost = 20       #How many energy units the mysis use up migrating

setGeneric( "nextTime", function(object, ...){standardGeneric("nextTime")})
setMethod("nextTime","mysis", 
          function(object, foodAvail, foodVar, time){       #Takes in the mysis object and the ratio of food quality at a given time. 
            
            #Run the draws: 
            migrationDraw   = runif(1) #random number between 0 and 1, this will be used for the migration decision
            migrationDraw_2 = runif(1) #draw for second migration decision, this time compared to logistic curve
            predationDraw   = runif(1) #" " to see if killed by predation
            #Lets set some constants real quick: 
            migrationRisk = 0  #We don't really need to simulate predation.
            stayRisk      = 0 
#             migrationRisk = 0.0001   #Chance of being eaten if migrating
#             stayRisk      = 0.00001  #Chance of being eaten if they stay on the bottom
#             
            #Energy rewards:
            migrationReward = rnorm(1, (rewardUnits*(1 + foodAvail)), foodVar)
            stayReward      = migrationReward * .2
            threshold = 0.2  #Let's make sure that the benthic reward can never drop below a threshold.
            if(stayReward < threshold){
              stayReward = threshold
            }
            
            sunset  = 18 #hardcode sunset, fill this with real data later. 
            sunrise = 7
            conditionCurve = migrationProb(object@energy)

            if (object@energy <= 0){ #did the mysis starve?
              object@alive = FALSE
              object@energy = 0
            }     
            #if the mysis is in good condition use standard migration chance, if bad use 1 - chance
            if ((time %% 24 == sunset + 1)&&(object@alive)){
              if (migrationDraw < foodAvail) {
                object@migrating = migrationDraw_2 > conditionCurve 
                object@energy = object@energy - migrationCost 
              } else {
                 object@migrating = FALSE
              } 
            } else if (time %% 24 == sunrise){ #bring them back down at sunrise. 
              object@migrating = FALSE
            }    
          
            #Here is the decision tree:
            if (object@alive){ #If the mysis is alive let's run the decision tree
              if (object@migrating){ 
                object@depth = depthData[time] #Grab the mysocline limit at this hour   
                
                if (predationDraw > migrationRisk){ #The mysis evades predation
                  object@energy = object@energy + migrationReward 
                } else { #Mysis is eaten
                  object@alive = FALSE
                } 
              } else { #The mysis didn't migrate
                object@depth = 100   
                
                if (predationDraw > stayRisk){ #The mysis evades predation
                  object@energy = object@energy +  stayReward                 
                } else { #Mysis is eaten
                  object@alive = FALSE
                }
              }
            }
            object} #Return the mysis
          )
```


### Testing it: 

Now that the model is all set up can run it over a small subset of the eventual numbers (~10,000 mysids over 8,700 hours) and see what results we get. 

```{r}
mysids = NULL
numOfMysids = 250
mysisNames = paste("mysis",(seq(1,numOfMysids)), sep = "")

for (i in 1:numOfMysids){
  initialenergy = rnorm(1,150,25) #draw initial energy from normal dist centered at 30 with stdev of 5. 
  mysids = c(mysids, new("mysis", energy = initialenergy) )
}
```

... and now we can run the model on those mysids: 

```{r}
migrations = NULL #Initialize some dataframes to keep track of the migrations and conditions values for the mysids. 
conditions = NULL
alive      = NULL
hours = 1:(24*365) #150 days
for (mysid in mysids){ #loop through the mysis
  migration = NULL  
  condition = NULL
  for (i in hours){
    mysid    = nextTime(mysid, foodAvail[i], foodVar[i] , i)
    migration = c(migration, mysid@depth)
    condition = c(condition, mysid@energy)
    if (i == length(hours)) { alive = c(alive, mysid@alive)}
  }
  migrations = cbind(migrations, migration) # add to the object of migrations. 
  conditions = cbind(conditions, condition)
} 

```

## What do we have?

Let's plot this small run to see what kind of behavior we are getting. 

```{r echo=FALSE, results='hide'}
migrations = cbind(hours,migrations)

#Make the dataframe for the migrations
migrations_df = as.data.frame(migrations)
names(migrations_df) = c("hours", mysisNames)
migrations_plot = melt(migrations_df, id = 'hours')

#Render the plot for the migrations
# a = ggplot(migrations_plot, aes(x = hours, y = -value, group = variable)) + geom_line(aes(color = variable)) + 
#   labs(title = "Mysis migration patterns for first 100 days of year", y = "Depth below surface", x = "") + theme(legend.position="none")

#Assemble the dataframe for condition
conditions_df = as.data.frame(cbind(hours, conditions))
names(conditions_df) = c("hours", mysisNames)
conditions_plot = melt(conditions_df, id = 'hours')

#Find portion who died
mysidsWhoDied = tail(conditions_df, 1)[1:numOfMysids+1] <= 0
mortalityRate = sum(mysidsWhoDied)/numOfMysids

subtitleText = paste(numOfMysids, "mysids, Reward:", rewardUnits, "Migration Cost:", migrationCost, ", Mortality Rate:",mortalityRate)

#Draw condition plot
b = ggplot(conditions_plot, aes(x = hours, y = value, group= variable)) + stat_smooth(aes(color = variable),method = "gam", formula = y ~ s(x, bs = "cs")) + 
  labs(title = "blah!", y = "energy units") + theme(legend.position="none") +  ggtitle(bquote(atop(.("Mysis Condition Over Year"), atop(italic(.(subtitleText)), "")))) 

b
ggsave(file = paste("figures/condition", rewardUnits, "_", migrationCost,".pdf", sep = ""))
# #Plot both migrations and condition on top of eachother. 
# Plot = arrangeGrob(a,b,ncol=1, main = paste("Migration and Condition,", "Mortality Rate:", mortalityRate) )
# Plot 


```

## What are the patterns of the mysis that survived?

By plotting a histogram of the days migrated by surviving mysids we can get a glimps into the more succesful strategies employed. 

First we make a function that takes in migration height data and returns how many days the mysid migrated:
```{r}
howManyMigrations = function(mysid){return(365 - sum(mysid[seq(19,(365*24), 24)] == 100))}
```

Next, we run this through the mysids that survived. 
```{r}
mysidsWhoLived = mysidsWhoDied == FALSE

daysMigrated = NULL

for(indv in mysisNames[mysidsWhoLived]){ 
  daysSurvived = howManyMigrations(migrations_df[[indv]])
  daysMigrated = c(daysMigrated, daysSurvived)
  }

survivingMysis = data.frame(daysMigrated)

ggplot(survivingMysis, aes(x = daysMigrated)) + 
  geom_histogram(aes(y=..count../sum(..count..)), color = "black", fill = "#7fc97f", binwidth = 5) + 
  labs(title = "Days Migrated for Surviving Mysids", x = "Days Migrated", y = "Freq") 

ggsave(paste("figures/days_migrated", rewardUnits, "_", migrationCost,".pdf", sep = ""))
```
