---
title: "Main Model"
author: "Nick Strayer"
date: "March 8, 2015"
output:
  html_document:
    fig_height: 6
    fig_width: 9
---

```{r echo=FALSE}
library(ggplot2)
setwd("/Users/Nick/mysisModeling") #Doesnt actually do anything in chunks.
```
---

## Check for the existance of needed datafiles. 

If the files are not present, then run their corresponding scripts and generate them. 

The corresponding R Markdown scripts associated with these files can be found on RPub: 

  - [Solar Data Cleaning](http://rpubs.com/nstrayer/64339)
  
  - [Thermocline Model](http://rpubs.com/nstrayer/thermoclineModel)

  - [Mysocline Depth Model](http://rpubs.com/nstrayer/64310)

```{r}
setwd("/Users/Nick/mysisModeling") #Locate ourselves in the computer. 

#Check for solar data
if(!file.exists("data/light_day_moon_hour.csv")){ source("dataGen/solarData.r") } 

#Check for thermocline data
if(!file.exists("data/Depth_Thermocline_Hour.csv")){ source("dataGen/thermoclineLevels.r") } 

#Mysocline data
if(!file.exists("data/mysocline_hour.csv")){ source("dataGen/depthModel.r") }

#Mysocline data
if(!file.exists("data/FoodAvail_Hour.csv")){ source("dataGen/foodAvailability.r") }

#Read in the data now. 
depthData = read.csv("data/mysocline_hour.csv")$x
foodAvail = read.csv("data/FoodAvail_hour.csv")$foodAvail
```

## Class and method declarations: 

I have chosen to use the S4 object oriented programing style. The details are minor but more info can be found [here](https://www.stat.auckland.ac.nz/~stat782/downloads/08-Objects.pdf). 

### Set up the mysis class: 

This basically is a digital mysis that simplistically holds a couple of peices of information and upon which the model will act. 

```{r}
setClass("mysis",
         representation(
           energy    = "numeric",    # The energy reserves are a numeric value
           migrating = "logical",    # If mysis is migrating is a logical value
           alive     = "logical",    # Alive still?
           depth     = "numeric"),   
         prototype(
           energy    = 0,            # The default instantiated mysis starts with zero energy...
           migrating = FALSE,        # and not migrating...
           alive     = TRUE,         # alive
           depth     = 100),         # and at the bottom.  
         )

#Set the show method, basically how we want the program to display the info about the mysis on calling. 
setMethod("show", "mysis", 
          function(object){
            if (object@alive){
              print(object@energy)
              print(object@migrating)
              print(object@depth)
            } else {
              print("dead")
            }
          })

```

### The decision tree model: 

This is a stochastic model, meaning it utilizes randomness to simulate the real world. 

The function of the decision tree is such: 

- Draw random number to decide if migrating: 
  - If the number drawn is above the ratio needed to migrate (food ratio) then the mysis migrates. 
    - After migration is decided, another draw is done to see if the mysis is killed by predation. 
 
 
![model tree](../figures/modelTree.pdf) 

```{r}
setGeneric( "nextTime", function(object, ...){standardGeneric("nextTime")})
setMethod("nextTime","mysis", 
          function(object, foodRatio, time){       #Takes in the mysis object and the ratio of food quality at a given time. 
            
            #Run the draws: 
            migrationDraw = runif(1) #random number between 0 and 1, this will be used for the migration decision
            predationDraw = runif(1) #" " to see if killed by predation
            
            #Lets set some constants real quick: 
            migrationRisk = 0.0001      #Chance of being eaten if migrating
            stayRisk      = 0.00001     #Chance of being eaten if they stay on the bottom
            
            #Energy rewards:
            costToMigrate   = 5
            rewardUnits     = 10
            migrationReward = rnorm(1,rewardUnits, rewardUnits*.5) - costToMigrate
            stayReward      =  rnorm(1,(rewardUnits * ( 1 - foodRatio)), rewardUnits*.5) 
            #this makes the rewards responsive to conditions
            
            #Migration Thresholds: 
            energyThreshold = 100;  #totally arbitrary. Above is high cond, below is low. 
            goodCondition  = object@energy > energyThreshold #will return true if high condition, false if low. 
            
            if (goodCondition){ #if the mysis is in good condition use standard migration chance, if bad use 1 - chance
              object@migrating = migrationDraw < foodRatio
            } else {
              object@migrating = migrationDraw < ( 1 - foodRatio)
            }
            
            #Here is the decision tree:
            if (object@alive){ #If the mysis is alive let's run the decision tree
  
              if (object@migrating){ 
                object@depth = depthData[time] #Grab the mysocline limit at this hour 
                
                if (predationDraw > migrationRisk){ #The mysis evades predation
                  object@energy = object@energy + migrationReward 
                
                } else { #Mysis is eaten
                  object@alive = FALSE
                }
                
              } else { #The mysis didn't migrate
                
                object@depth = 100
                
                if (predationDraw > stayRisk){ #The mysis evades predation
                  object@energy = object@energy +  stayReward
                  
                } else { #Mysis is eaten
                  object@alive = FALSE
                }
              }
            }
            object} #Return the mysis
          )

```


### Testing it: 

Now that the model is all set up can run it over a small subset of the eventual numbers (~10,000 mysids over 8,700 hours) and see what results we get. 

```{r}
mysids = NULL
numOfMysids = 5

for (i in 1:numOfMysids){
  initialenergy = sample(5:30,1) #draw initial energy from normal dist centered at 30 with stdev of 5. 
  mysids = c(mysids, new("mysis", energy = initialenergy) )
}
```

... and now we can run the model on those mysids: 

```{r}
migrations = NULL
hours = 1:(24*150) #150 days
for (mysid in mysids){ #loop through the mysis
  migration = NULL 
  for (i in hours){
#     foodMean = foodAvail$foodAvail
#     foodVar  = foodAvail$variability
    mysid    = nextTime(mysid, foodAvail[i], i)
    migration = c(migration, mysid@depth)
  }
  migrations = cbind(migrations, migration) # add to the object of migrations. 
} 
```

## What do we have?

Let's plot this small run to see what kind of behavior we are getting. 

```{r echo=FALSE}
migrations = cbind(migrations, hours)

migrationsDf = as.data.frame(migrations)
colnames(migrationsDf) = c("mysis1", "mysis2", "mysis3", "mysis4", "mysis5","hours")
ggplot(migrationsDf, aes(y = -mysis1, x = hours)) + geom_line() + 
  labs(title = "Mysis migration patterns for first 10 days of year", y = "Depth below surface")
```